\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{parskip}
\usepackage{times}
\setlength{\parindent}{0pt}

% Defining the document title and headers
\title{State Universities Offering Electrical Engineering Programs in the United States}
\author{}
\date{June 2025}

\begin{document}

\maketitle

\section*{Introduction}
This document provides a partial list of state universities in the United States that offer electrical engineering programs, based on available information from web sources. The list is not exhaustive due to the lack of a comprehensive, up-to-date dataset. For complete and accurate information, please consult official university websites or accreditation bodies such as ABET (Accreditation Board for Engineering and Technology).

\section*{List of State Universities with Electrical Engineering Programs}
The following table lists selected state universities known to offer electrical engineering programs, organized alphabetically by state. Note that program offerings may vary, and verification with each institution is recommended.

\begin{longtable}{p{0.3\textwidth} p{0.2\textwidth} p{0.4\textwidth}}
\toprule
\textbf{University Name} & \textbf{State} & \textbf{Program Notes} \\
\midrule
\endhead
University of Alabama & Alabama & Offers BS, MS, and PhD in Electrical Engineering, with focus areas like power systems and electronics. \\
Auburn University & Alabama & ABET-accredited BS in Electrical Engineering; strong emphasis on control systems and communications. \\
University of Arizona & Arizona & BS, MS, and PhD programs; research in microelectronics and renewable energy. \\
Arizona State University & Arizona & Highly ranked online and on-campus BS in Electrical Engineering; focus on energy systems and robotics. \\
University of Arkansas & Arkansas & BS and MS programs; research in power electronics and signal processing. \\
University of California, Berkeley & California & Top-ranked globally; offers BS, MS, and PhD with specializations like integrated circuits. \\
University of California, Los Angeles & California & ABET-accredited BS; strong graduate programs in communications and embedded systems. \\
Colorado State University & Colorado & BS, MS, and PhD; research in electromagnetics and energy systems. \\
University of Colorado Boulder & Colorado & Offers BS and graduate degrees; focus on photonics and power electronics. \\
University of Connecticut & Connecticut & BS, MS, and PhD; research in control systems and nanotechnology. \\
University of Delaware & Delaware & BS and graduate programs; emphasis on signal processing and materials. \\
University of Florida & Florida & ABET-accredited BS; strong research in semiconductors and wireless communications. \\
Florida State University & Florida & BS in Electrical Engineering; collaborative programs with FAMU. \\
University of Georgia & Georgia & BS and MS; growing program in electronics and automation. \\
Georgia Institute of Technology & Georgia & One of the largest ECE programs globally; BS, MS, PhD with AI and robotics focus. \\
University of Hawaii at Manoa & Hawaii & BS and graduate degrees; research in telecommunications and renewable energy. \\
University of Illinois Urbana-Champaign & Illinois & Top-ranked; BSBai MS, and PhD in areas like microelectronics and machine learning. \\
Iowa State University & Iowa & ABET-accredited BS; research in power systems and VLSI design. \\
University of Iowa & Kentucky & BS, MS, and PhD; focus on control systems and biomedical engineering. \\
University of Kentucky & Kentucky & BS and graduate programs; emphasis on power and energy systems. \\
Louisiana State University & Louisiana & BS, MS, and PhD; research in electronics and telecommunications. \\
University of Maryland, College Park & Maryland & Highly ranked BS and graduate programs; focus on quantum engineering and communications. \\
University of Massachusetts Amherst & Massachusetts & BS, MS, and PhD; research in RF systems and embedded computing. \\
Michigan State University & Michigan & BS and graduate degrees; strong in power systems and controls. \\
University of Michigan & Michigan & Top-ranked; BS, MS, and PhD with focus on autonomous systems and photonics. \\
University of Minnesota & Minnesota & BS, MS, and PhD; research in microsystems and energy conversion. \\
Mississippi State University & Mississippi & BS and graduate programs; emphasis on power systems and high-voltage engineering. \\
University of Missouri & Missouri & BS, MS, and PhD; research in electronics and signal processing. \\
University of Nebraska-Lincoln & Nebraska & ABET-accredited BS; focus on communications and robotics. \\
University of Nevada, Reno & Nevada & BS and graduate degrees; research in renewable energy and sensors. \\
University of New Hampshire & New Hampshire & BS in Electrical Engineering; emphasis on embedded systems. \\
Rutgers University & New Jersey & BS, MS, and PhD; research in wireless networks and machine learning. \\
University of New Mexico & New Mexico & BS and graduate programs; focus on optics and semiconductor devices. \\
University at Buffalo & New York & BS, MS, and PhD; research in power electronics and nanotechnology. \\
North Carolina State University & North Carolina & Highly ranked BS and graduate programs; focus on power systems and IoT. \\
University of North Dakota & North Dakota & BS in Electrical Engineering; emphasis on energy systems. \\
Ohio State University & Ohio & BS, MS, and PhD; research in electromagnetics and autonomous systems. \\
Oklahoma State University & Oklahoma & BS and graduate degrees; focus on power and control systems. \\
University of Oklahoma & Oklahoma & BS, MS, and PhD; research in radar and telecommunications. \\
Oregon State University & Oregon & BS and graduate programs; strong in energy systems and microelectronics. \\
Penn State University & Pennsylvania & ABET-accredited BS; research in RF systems and photonics. \\
University of Rhode Island & Rhode Island & BS in Electrical Engineering; focus on signal processing and robotics. \\
University of South Carolina & South Carolina & BS, MS, and PhD; research in power electronics and controls. \\
South Dakota State University & South Dakota & BS in Electrical Engineering; emphasis on rural power systems. \\
University of Tennessee & Tennessee & BS, MS, and PhD; research in power systems and embedded computing. \\
Texas A\&M University & Texas & Top-ranked; BS, MS, and PhD with focus on power and analog circuits. \\
University of Texas at Austin & Texas & Highly ranked BS and graduate programs; research in integrated circuits and energy. \\
Utah State University & Utah & BS and graduate degrees; focus on space systems and controls. \\
University of Vermont & Vermont & BS in Electrical Engineering; emphasis on sustainable energy. \\
Virginia Tech & Virginia & BS, MS, and PhD; research in power electronics and wireless systems. \\
University of Washington & Washington & Top-ranked; BS, MS, and PhD with focus on embedded systems and AI. \\
West Virginia University & West Virginia & BS and graduate programs; research in power systems and biometrics. \\
University of Wisconsin-Madison & Wisconsin & BS, MS, and PhD; strong in control systems and microelectronics. \\
University of Wyoming & Wyoming & BS in Electrical Engineering; focus on energy and telecommunications. \\
\bottomrule
\caption{Selected state universities offering electrical engineering programs. This list is partial and should be verified with official sources.}
\label{tab:ee_programs}
\end{longtable}

\section*{Notes}
\begin{itemize}
    \item The universities listed are public institutions identified as state universities based on common classifications. Private institutions and non-state public colleges are excluded.
    \item Program details (e.g., BS, MS, PhD) are based on typical offerings for electrical engineering departments. Specific concentrations or accreditation status may vary.
    \item Sources include general knowledge and web information from rankings like US News, EduRank, and QS World University Rankings, but no single source provided a complete list of state universities with electrical engineering programs.
    \item For a comprehensive list, consider consulting ABET’s accredited program search (www.abet.org) or the National Center for Education Statistics (nces.ed.gov).
\end{itemize}

\end{document}